Partial Least Squares Regression with Multiple Functional and Scalar Predictors

### 2025-07-09 manuscript/main.tex
infinite expansion
from Yingying's paper

@@

For $k = 1, \ldots, K$, we project the $k$th functional predictor onto 
$\operatorname{span}(\Phi^{(k)})$:
\begin{equation*}
	X_i^{(k)} \approx \sum_{m=1}^M c_{im}^{(k)} \phi_m^{(k)}, \quad i = 1, \ldots, n,
\end{equation*}
where the coefficients $c_{im}^{(k)}$ satisfy
\begin{equation*}
	X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} \phi_m^{(k)}(t_k), \quad \text{for all } t_k \in \mathbf{t}_k \subset \mathcal{T}_k.
\end{equation*}
Thus, the $n$ observations of the $k$th functional predictor can be represented by an $n \times M$ matrix of coefficients, 	where rows index samples and columns index basis functions:
\begin{equation*}
		\widetilde{\mathbf{C}}^{(k)} := ( c_{im}^{(k)} ).
\end{equation*}

@@

	\subsection{Proof of Propositions \ref{Proposition: generalized rayleigh quotient} and \ref{Proposition: Regularized generalized rayleigh quotient} }\label{section:proof:eigenproblem_derivation}
	%
	%
	\begin{align*}
	\widehat{\textnormal{Cov}} (
	\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y 
	) 
	&= 
	\frac{1}{n} \sum \limits_{i=1}^n 
	 y_i
	\langle \widetilde{\mathbf{W}}_i, \boldsymbol{\xi} \rangle_\mathcal{H} \\
	&= 
	 \frac{1}{n} 
	 \sum \limits_{i=1}^n
	 y_i
	   \langle \widetilde{X_i}, \psi \rangle_\mathcal{F}  
	   +  
	   \frac{1}{n} 
	   \sum \limits_{i=1}^n  
	   y_i
	   \langle \mathbf{Z}_i, 
	   \boldsymbol{\theta} 
	   \rangle 
	    \\
	&=  \frac{1}{n} \sum \limits_{i=1}^n  \left\{ \sum \limits_{k=1}^K \int_{\mathcal{T}_k} X_i(t_k) \psi^{(k)}(t_k) dt_k \right\} Y_i +  \frac{1}{n} \sum \limits_{i=1}^n \mathbf{Z}_i^\top \boldsymbol{\theta} Y_i \\
	&=  \frac{1}{n} \sum \limits_{i=1}^n \left[ \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \left\{\sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t) \right\} \left\{\sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t) \right\} dt_k \right] Y_i + \frac{1}{n} \sum \limits_{i=1}^n \mathbf{Z}_i^\top \boldsymbol{\theta} Y_i \\
	&= n^{-1} \widetilde{\mathbf{y}}^\top \left[ \sum \limits_{k=1}^K \widetilde{\mathbf{C}}^{(k)} \left\{ \int_{\mathcal{T}_k} \mathbf{b}^{(k)}(t_k) \mathbf{b}^{(k)}(t_k)^\top dt_k \right\} \mathbf{a}^{(k)} \right] + n^{-1}\widetilde{\mathbf{y}}^\top \mathbf{Z} \boldsymbol{\theta} \\
	&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C} J \mathbf{a}) + n^{-1}\widetilde{\mathbf{y}}^\top \mathbf{Z} \boldsymbol{\theta} \\
	&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C}J\mathbf{a} + \mathbf{Z} \boldsymbol{\theta}).
\end{align*}

>>>>

Given an orthonormal basis $\{b_l(t)\}$, the functional predictors and the corresponding regression coefficients can be decomposed into
\begin{align*}
	X_{ij}(t) = \sum_{l=1}^\infty \theta_{ijl} \, b_l(t), \quad 
	\beta_j(t) = \sum_{l=1}^\infty \eta_{0,jl} \, b_l(t),
\end{align*}
where $\theta_{ijl}$ and $\eta_{0,jl}$ are the coefficients of $X_{ij}(t)$ and $\beta_j(t)$ corresponding to the $l$th basis function $b_l(t)$, respectively. While it is possible to use a different basis for each functional predictor, we adopt a common basis system for all predictors for notational simplicity.

@@

For $k = 1, \ldots, K$, we project the $k$th functional predictor $X_i^{(k)}$ onto 
$\operatorname{span}(\Phi^{(k)})$:
\begin{equation*}
	X_i^{(k)} \approx \sum_{m=1}^M c_{im}^{(k)} \phi_m^{(k)}, \quad i = 1, \ldots, n,
\end{equation*}
where the coefficients $c_{im}^{(k)}$ satisfy
\begin{equation*}
	X_i^{(k)}(t_k) = \sum_{m=1}^M c_{im}^{(k)} \phi_m^{(k)}(t_k), \quad \text{for all } t_k \in \mathbf{t}_k \subset \mathcal{T}_k.
\end{equation*}
Thus, the $n$ observations of the $k$th functional predictor can be represented by an $n \times M$ matrix of coefficients, 	where rows index samples and columns index basis functions:
\begin{equation*}
		\widetilde{\mathbf{C}}^{(k)} := ( c_{im}^{(k)} ).
\end{equation*}

@@

	\subsection{Proof of new Proposition}
		\begin{align*}
		\widehat{\textnormal{Cov}} (
		\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y 
		) 
		&= 
		\frac{1}{n} \sum \limits_{i=1}^n 
		y_i
		\langle \widetilde{\mathbf{W}}_i, \boldsymbol{\xi} \rangle_\mathcal{H} \\
		&= 
		\frac{1}{n} 
		\sum \limits_{i=1}^n
		y_i
		\langle \widetilde{X_i}, \psi \rangle_\mathcal{F}  
		+  
		\frac{1}{n} 
		\sum \limits_{i=1}^n  
		y_i
		\langle \mathbf{Z}_i, 
		\boldsymbol{\theta} 
		\rangle 
		\\
		&=  \frac{1}{n} \sum \limits_{i=1}^n  \left\{ \sum \limits_{k=1}^K \int_{\mathcal{T}_k} X_i(t_k) \psi^{(k)}(t_k) dt_k \right\} Y_i +  \frac{1}{n} \sum \limits_{i=1}^n \mathbf{Z}_i^\top \boldsymbol{\theta} Y_i \\
		&=  \frac{1}{n} \sum \limits_{i=1}^n \left[ \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \left\{\sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t) \right\} \left\{\sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t) \right\} dt_k \right] Y_i + \frac{1}{n} \sum \limits_{i=1}^n \mathbf{Z}_i^\top \boldsymbol{\theta} Y_i \\
		&= n^{-1} \widetilde{\mathbf{y}}^\top \left[ \sum \limits_{k=1}^K \widetilde{\mathbf{C}}^{(k)} \left\{ \int_{\mathcal{T}_k} \mathbf{b}^{(k)}(t_k) \mathbf{b}^{(k)}(t_k)^\top dt_k \right\} \mathbf{a}^{(k)} \right] + n^{-1}\widetilde{\mathbf{y}}^\top \mathbf{Z} \boldsymbol{\theta} \\
		&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C} J \mathbf{a}) + n^{-1}\widetilde{\mathbf{y}}^\top \mathbf{Z} \boldsymbol{\theta} \\
		&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C}J\mathbf{a} + \mathbf{Z} \boldsymbol{\theta}).
	\end{align*}	
	\subsection{Proof of Propositions \ref{Proposition: generalized rayleigh quotient} and \ref{Proposition: Regularized generalized rayleigh quotient} }\label{section:proof:eigenproblem_derivation}
	%
	%
	\begin{align*}
	\widehat{\textnormal{Cov}} (
	\langle \widetilde{\mathbf{W}}, \boldsymbol{\xi} \rangle_{\mathcal{H}}, Y 
	) 
	&= 
	\frac{1}{n} \sum \limits_{i=1}^n 
	 y_i
	\langle \widetilde{\mathbf{W}}_i, \boldsymbol{\xi} \rangle_\mathcal{H} \\
	&= 
	 \frac{1}{n} 
	 \sum \limits_{i=1}^n
	 y_i
	   \langle \widetilde{X_i}, \psi \rangle_\mathcal{F}  
	   +  
	   \frac{1}{n} 
	   \sum \limits_{i=1}^n  
	   y_i
	   \langle \mathbf{Z}_i, 
	   \boldsymbol{\theta} 
	   \rangle 
	    \\
	&=  \frac{1}{n} \sum \limits_{i=1}^n  \left\{ \sum \limits_{k=1}^K \int_{\mathcal{T}_k} X_i(t_k) \psi^{(k)}(t_k) dt_k \right\} Y_i +  \frac{1}{n} \sum \limits_{i=1}^n \mathbf{Z}_i^\top \boldsymbol{\theta} Y_i \\
	&=  \frac{1}{n} \sum \limits_{i=1}^n \left[ \sum \limits_{k=1}^K \int_{\mathcal{T}_k} \left\{\sum_{m=1}^M c_{im}^{(k)} b^{(k)}_{m}(t) \right\} \left\{\sum_{m=1}^M a^{(k)}_{m} b^{(k)}_m(t) \right\} dt_k \right] Y_i + \frac{1}{n} \sum \limits_{i=1}^n \mathbf{Z}_i^\top \boldsymbol{\theta} Y_i \\
	&= n^{-1} \widetilde{\mathbf{y}}^\top \left[ \sum \limits_{k=1}^K \widetilde{\mathbf{C}}^{(k)} \left\{ \int_{\mathcal{T}_k} \mathbf{b}^{(k)}(t_k) \mathbf{b}^{(k)}(t_k)^\top dt_k \right\} \mathbf{a}^{(k)} \right] + n^{-1}\widetilde{\mathbf{y}}^\top \mathbf{Z} \boldsymbol{\theta} \\
	&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C} J \mathbf{a}) + n^{-1}\widetilde{\mathbf{y}}^\top \mathbf{Z} \boldsymbol{\theta} \\
	&= n^{-1} \widetilde{\mathbf{y}}^\top (\widetilde{C}J\mathbf{a} + \mathbf{Z} \boldsymbol{\theta}).
\end{align*}
